One page description

### 2016-12-07 publications/Journal/3-features.tex
Update design section and UML to switch from core to kernel if referring to the simulator instance.
Signed-off-by: Ben Cardoen <bencardoen@gmail.com>

We will not make a detailed comparison with \textit{PythonPDEVS} here, but only mention some supported features.
\textit{Dxex} supports, similarly to \textit{PythonPDEVS}, the following features: direct connection~\cite{SymbolicFlattening}, \textsf{Dynamic Structure DEVS}~\cite{DSDEVS}, termination conditions~\cite{JDF}, and a modular tracing and scheduling framework~\cite{PythonPDEVS}.
We do not elaborate on these features in this paper.
But whereas \textit{PythonPDEVS} only supports optimistic synchronization, \textit{dxex} support multiple synchronization protocols (though only in parallel).
This is in line with the design principle of \textit{PythonPDEVS}: allow users to pass performance hints~\cite{PythonPDEVS_ACTIMS} to the simulation kernel.
In our case, a user can configure the simulation kernel with the synchronization protocol to use, or even switch the synchronization protocol during runtime.
Our implementation in C++11 also allows for (compiled) optimizations which were plainly impossible in an interpreted language.
\textit{Dxex} uses new optimizations from the C++14 standard when possible.
The C++11 standard threading primitives are used to run the different simulation cores.
Within a single simulation core, OpenMP~\cite{openmp4} is used to parallelize the transition functions, as is usual in \textsf{Parallel DEVS}.

@@

\subsection{Synchronization protocols}
We previously explained the existence of different synchronization protocols, each optimized for a specific kind of model.
As no single synchronization protocol is ideal for all domains, a general purpose simulation tool should support multiple protocols.
We argue that a general purpose simulation tool should support single core, conservative, and optimistic synchronization, as is the case for \textit{dxex}. Futhermore it should be possible to combine each inter core synchronization protocol with the parallel execution of events defined by \pSim.

@@

Data exchange between different simulation cores happens through shared memory, using the new C++11 synchronization primitives.
Whereas this was also possible in previous versions of the C++ standard, by falling back to non-portable C functions, it was not a part of the C++ language standard.
C++11 further allows us to make the implementation portable, as well as more efficient: the compiler makes further optimizations to heavily used components.

@@

An overview of \textit{dxex}'s design is given in Figure~\ref{fig:class_diagram}.
It shows that there is a simulation \texttt{Core}, which simulates the \texttt{AtomicModel}s connected to it.
The superclass \texttt{Core} represents the single core simulation kernel.
Subclasses define specific variants, such as \texttt{ConservativeCore} (conservative synchronization), \texttt{OptimisticCore} (optimistic synchronization), and \texttt{DynamicCore} (\textsf{Dynamic Structure DEVS}).

@@

% protocol or inherent parallelism?
\subsubsection{\pSim}
In our tool each simulation core, regardless of inter core synchronization, is capable of executing concurrent transitions in parallel. A configurable number of threads is allocated to optimally divide the processing load of concurrent transitions. The parallel execution of transitions introduces some overhead in thread pooling and locking, and disallows some optimizations in optimally rescheduling models. The concurrency of events and the combined computational load of the transitions has to outweigh this overhead to obtain a significant speedup, as we will demonstrate in Section~\textsc{\nameref{sec:4-performance}}.

>>>>

We will not make a detailed comparison with \textit{PythonPDEVS} here, but only mention some supported features.
\textit{Dxex} supports, similarly to \textit{PythonPDEVS}, the following features: direct connection~\cite{SymbolicFlattening}, \textsf{Dynamic Structure DEVS}~\cite{DSDEVS}, termination conditions~\cite{JDF}, and a modular tracing and scheduling framework~\cite{PythonPDEVS}.
We do not elaborate on these features in this paper.
But whereas \textit{PythonPDEVS} only supports optimistic synchronization, \textit{dxex} support multiple synchronization protocols (though only in parallel).
This is in line with the design principle of \textit{PythonPDEVS}: allow users to pass performance hints~\cite{PythonPDEVS_ACTIMS} to the simulation kernel.
In our case, a user can configure the simulation kernel with the synchronization protocol to use, or even switch the synchronization protocol during runtime.
Our implementation in C++11 also allows for (compiled) optimizations which were plainly impossible in an interpreted language.
\textit{Dxex} uses new optimizations from the C++14 standard when possible.
The C++11 standard threading primitives are used to run the different simulation kernels.
Within a single simulation kernel, OpenMP~\cite{openmp4} is used to parallelize the transition functions, as is usual in \textsf{Parallel DEVS}.

@@

\subsection{Synchronization protocols}
We previously explained the existence of different synchronization protocols, each optimized for a specific kind of model.
As no single synchronization protocol is ideal for all domains, a general purpose simulation tool should support multiple protocols.
We argue that a general purpose simulation tool should support single core, conservative, and optimistic synchronization, as is the case for \textit{dxex}. Futhermore it should be possible to combine each inter kernel synchronization protocol with the parallel execution of events defined by \pSim.

@@

Data exchange between different simulation kernels happens through shared memory, using the new C++11 synchronization primitives.
Whereas this was also possible in previous versions of the C++ standard, by falling back to non-portable C functions, it was not a part of the C++ language standard.
C++11 further allows us to make the implementation portable, as well as more efficient: the compiler makes further optimizations to heavily used components.

@@

An overview of \textit{dxex}'s design is given in Figure~\ref{fig:class_diagram}.
It shows that there is a simulation \texttt{Kernel}, which simulates the \texttt{AtomicModel}s connected to it.
The superclass \texttt{Kernel} represents the single core simulation kernel.
Subclasses define specific variants, such as \texttt{ConservativeKernel} (conservative synchronization), \texttt{OptimisticKernel} (optimistic synchronization), and \texttt{DynamicKernel} (\textsf{Dynamic Structure DEVS}).

@@

% protocol or inherent parallelism?
\subsubsection{\pSim}
In our tool each simulation kernel, regardless of inter kernel synchronization, is capable of executing concurrent transitions in parallel. A configurable number of threads is allocated to optimally divide the processing load of concurrent transitions. The parallel execution of transitions introduces some overhead in thread pooling and locking, and disallows some optimizations in optimally rescheduling models. The concurrency of events and the combined computational load of the transitions has to outweigh this overhead to obtain a significant speedup, as we will demonstrate in Section~\textsc{\nameref{sec:4-performance}}.
